# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

Load data from zip-file.

```{r}
data <- read.csv(unz("activity.zip", "activity.csv"), header=T, sep=",")
```

## What is mean total number of steps taken per day?

Creating boxplot (histogram) of total steps per day.

```{r}
StepsPerDayPlot <- function(par_data, plot_color)
{
    diff_dates=unique(par_data$date)
    steps_by_day=rep.int(0,length(diff_dates))
    iter=1:length(diff_dates)
    for(i in iter)
        steps_by_day[i]=sum(par_data[par_data$date==diff_dates[i],"steps"],na.rm=TRUE)
    barplot(steps_by_day,names.arg=diff_dates,col=plot_color)
    return(steps_by_day)
}
steps_by_day=StepsPerDayPlot(data,"darkgreen")
```

Calculating mean.

```{r}
mean(steps_by_day)
```

Calculating median.

```{r}
median(steps_by_day)
```

## What is the average daily activity pattern?

Creating time series plot.

```{r}
intervals=unique(data$interval)
interval_steps_over_days=rep(0,length(intervals))
interval_iter=1:length(intervals)
for(i in interval_iter)
    interval_steps_over_days[i]=mean(data[data$interval==intervals[i],"steps"],na.rm=TRUE)
plot(intervals, interval_steps_over_days, type="l", col="darkblue")
```

Find 5-minute interval with maximum number of steps on average across all the days in the dataset.

```{r}
max_interval_index=which(interval_steps_over_days==max(interval_steps_over_days))
print(intervals[max_interval_index])
```

## Imputing missing values

Get total number of NA values in dataset.

```{r}
print(sum(is.na(data)))
```

The only column where the NA values are is "steps".

```{r}
print(sum(is.na(data$steps)))
```

Changing NA cells to average number of steps at corresponding **day** is not wise - 
number of steps have much variability depending on time interval. So it is more appropriate
to take mean value of steps during corresponding **time interval** across all days. Plus, such
averages are already calculated.

First, dublicate dataset.

```{r}
data2=data
```

Now, changing values.

```{r}
rows_with_NA=which(is.na(data2$steps))
for(i in rows_with_NA)
{
    mean_index=which(intervals==data2$interval[i])
    data2$steps[i]=interval_steps_over_days[mean_index]
}
```

Checking absense of NA values.

```{r}
print(sum(is.na(data2)))
```

Creating two boxplots (histograms) of total steps per day with original and 
modified datasets for easier comparison. Blue horizontal line is at the level of mean 
in both cases.

```{r}
par(mfrow=c(2,1))
steps_by_day_old=StepsPerDayPlot(data,"darkgreen")
abline(h=mean(steps_by_day_old),col="blue")
steps_by_day_new=StepsPerDayPlot(data2,"green")
abline(h=mean(steps_by_day_new),col="blue")
par(mfrow=c(1,1))
```

Old mean and median.

```{r}
mean(steps_by_day_old)
median(steps_by_day_old)
```

New mean and median.

```{r}
mean(steps_by_day_new)
median(steps_by_day_new)
```

For new dataset **mean** is higher which should be (we increased number of steps at days where
NA values were). The **median** is also higher (which not nesessarely *should be*, but it is so in this case).

## Are there differences in activity patterns between weekdays and weekends?

Creating new column for modified dataset - date_POSIXlt for dates in POSIXlt format. We use "GMT" timezone
through it is irrelevant for current analysis.

```{r}
data2$date_POSIXlt=strptime(data2$date,format="%Y-%m-%d",tz="GMT")
class(data2$date_POSIXlt)[1]
```

