# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Load data from zip-file.
```{r}
data <- read.csv(unz("activity.zip", "activity.csv"), header=T, sep=",")
```

Creating new column - date_POSIXlt for dates in POSIXlt format. We use "GMT" timezone
through it is irrelevant for current analysis.
```{r}
data$date_POSIXlt=strptime(data$date,format="%Y-%m-%d",tz="GMT")
```

## What is mean total number of steps taken per day?

```{r}
diff_dates=unique(data$date)
steps_by_day=rep.int(0,length(diff_dates))
iter=1:length(diff_dates)

for(i in iter)
    steps_by_day[i]=sum(data[data$date==diff_dates[i],"steps"],na.rm=TRUE)

barplot(steps_by_day,names.arg=diff_dates,col="darkgreen")
```
Mean total number of steps taken per day is `r mean(steps_by_day)`. 
Median is `r format(median(steps_by_day),scientific=FALSE)`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
